---
title: "clairezh-AN588-replication"
author: "claire zhang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading in the data:
```{r}
f <- "C:/Users/clair/Downloads/DATASET_CARDOSO&OTTONI.csv.csv"
d <- read.csv(f, header = TRUE, stringsAsFactors = FALSE)
head(d) # looks fine
```

## descriptive statistics:
authors stated that the SCNP capuchins visited the boxes at a mean = 213 visits/day; SD = 55; total individual visits = 1067. the FBV capuchins visited the boxes at a mean = 29 visits/day; SD = 12; total individual visits = 376

partitioning the two populations into separate datasets
```{r}
SCNP <- subset(d, POPULATION == "SCNP")
FBV <- subset(d, POPULATION == "FBV")
SCNP
FBV # looks fine!
```

calculating mean and SD number of visits for the SCNP population:
```{r}
sum(SCNP$NUMVISIT)
mean(SCNP$NUMVISIT)
sd(SCNP$NUMVISIT)
```

the total number of visits adds up fine. maybe they made a typo? I don't understand how the dataset I have could possibly turn out a mean of 213 visits/days of presentation.

looking at the FBV population:
```{r}
sum(FBV$NUMVISIT)
mean(FBV$NUMVISIT)
sd(FBV$NUMVISIT)
```

okay. well the total doesn't even add up correctly, much less the mean and standard deviation hello? what am I doing wrong, because there's no way THEIR data is just crazy incorrect.

trying the "Mann-Whitney test" they do. They say the Z = -4541, p < 0.0001. I'm honestly not even sure what they're calculating the statistics *of*, but my best guess is that it's comparing the frequency of probe use in the SCNP group vs. the FBV population. also, I'm assuming that the Mann-Whitney test is an unpaired two-sample Wilcoxon test

```{r}
mw.probes <- wilcox.test(SCNP$PROBE.EVENT, FBV$PROBE.EVENT)
mw.probes
```

this doesn't give me enough information, except that the p-value isn't correct if I just use the wilcox.test() function. therefore, I'm going to try calculating it by hand:
